---
title: 'Assignment 7.1 : Fit a Logistic Regression Model to the Thoracic Surgery Binary Dataset'
author: "Ragunath Gunasekaran"
date: "July 21 2020"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---


## Footnote
This is a Footnote test.

## Citations

* R for Everyone
* Discovering Statistics Using R

```{r dataframe code, include=FALSE}
library(foreign)
surgery_df <- read.arff("ThoraricSurgery.arff")
```


##  Exploratory Analysis

```{r plot, echo=FALSE, message=FALSE}
library(ggplot2)
plot(surgery_df[,1:8])
plot(surgery_df[,9:17])
ggplot(surgery_df, mapping = aes(x = DGN, y = PRE14)) + geom_point(mapping = aes(colour = Risk1Yr)) + xlab("Weakness before Surgery") + ylab("Diagnosis") 

```

##  Binary logistic regression model to the data set 
Training and Testing Data

Here I created data to train my models on adn then testing data to test the trained models on to see how well they work on data it was not trained on.

Fitted a binary logistic regression model to the data set that predicts whether or not the patient survived for one year (the Risk1Y variable) after the surger and going to use  Use the glm() function to perform the logistic regression.Also using a summary using the summary() function in our results

```{r plot1, echo=FALSE, message=FALSE}
TFRAC = 0.25
index = sample(1:nrow(surgery_df), TFRAC * nrow(surgery_df))
test.data <- surgery_df[index,]
train.data <- surgery_df[-index,]

surgery_df.glm <- glm(Risk1Yr ~ ., data = train.data, family = binomial)
summary(surgery_df.glm)

```


Computing the accuracy of your model, use the dataset to predict the outcome variable. The percent of correct predictions is the accuracy of your model

```{r plot3, echo=FALSE, message=FALSE}
library(MASS)
acc_vector <- rep(NA,9)
cutoff <- c(.1, .15, .2,.25, .3, .35, .4, .45,.5)
for(i in 1:9){
  predictions <- ifelse(predict(surgery_df.glm, test.data, type = "response")>cutoff[i],1,0)
  prob <- (predict(surgery_df.glm, test.data, type = "response"))
  pred.mod <- predictions
  pred.mod[which(pred.mod == 0)] <- "F"
  pred.mod[which(pred.mod == 1)] <- "T"
  acc_vector[i] <- sum(diag(prop.table(table(predictions, test.data$Risk1Yr))))}
show(acc_vector)
```

## Conclusion
From my the models that I made for this data set, the highest accuracy I acheived was about ~80%.



# References
1.  Chapter 8 from Discovering Statistics Using R.
2.  Understanding the Bias-Variance Tradeoff
3.  Calculating UAC: The Area Under a ROC Curve
4.  https://rpubs.com/melonda/190569
