---
title: 'Assignment 7.2 Assignment: Fit a Logistic Regression Model to Previous Dataset '
author: "Ragunath Gunasekaran"
date: "Aug 4 2020"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---


## Footnote
This is a Footnote test.

## Citations

* R for Everyone
* Discovering Statistics Using R

## Data Load - Data from CSV file load into Dataframe

```{r dataframe code, include=FALSE}
BinaryClassifier_df <- read.csv("binary-classifier-data.csv")
show(BinaryClassifier_df)
```


```{r dataframe StructureBin, echo=FALSE}
# Compactly Display The Structure Of An Arbitrary R Object -- Source 6
str(BinaryClassifier_df) 
```

Three variables x,y Labels
##  Logistic Regression with generalized linear models


a.The accuracy of the logistic regression classifier

# Exploring Data - Generating two models with x and y variable
```{r dataframe modelBinaryClassifier, include=FALSE}
modelBinaryClassifier <- glm(label ~ x + y, data = BinaryClassifier_df, family = "binomial")
```

## Summar of the model data for variable x and y
```{r dataframe summarydf, include=FALSE}
summary(modelBinaryClassifier)
```

#install.packages("ROCR")
#library(ROCR)


```{r dataframe logisticbindata, echo=FALSE}
# Calculate the Accuracy of the model ( Source 7 - explained about diabeties related model and Source 3)
#predict the binary classifer data set using type as Response and Variable

library(ROCR)
predict_modelBinaryClassifier <- predict(modelBinaryClassifier, newdata=BinaryClassifier_df, type="response")
predict_modeldataframelabel <- prediction(predict_modelBinaryClassifier, BinaryClassifier_df$label)

accuracyofmodel <- performance(predict_modeldataframelabel, measure = "auc")
accuracyofmodel <- accuracyofmodel@y.values[[1]]
accuracyofmodel

```

The accuracy of the logistic regression classifier for Binary Dataset is ~ 57%


## The accuracy of Nearest neighbors algorithm (KNN mode)
#source 6

## The accuracy of the logistic regression classifier and nearest neighbors
# Referred from Source 6

```{r dataframe knndataset, echo=FALSE}
library(class)
nor <-function(x) { (x -min(x))/(max(x)-min(x))   }
BinaryClassifier_dfknn <- as.data.frame(lapply(BinaryClassifier_df[,1:1],nor))
set.seed(12)

#selection with random of 75% data.
BinaryClassifier_dfknn1 <- sample(1:nrow(BinaryClassifier_df),size=nrow(BinaryClassifier_df)*0.75,replace = FALSE) 


#75% dataset - training
dt_Training1 <- BinaryClassifier_dfknn[BinaryClassifier_dfknn1,]

#25% data - testing
dt_testsample1 <- BinaryClassifier_dfknn[-BinaryClassifier_dfknn1,]


#75% dataset - training for variable y
dt_Training2 <- BinaryClassifier_dfknn[BinaryClassifier_dfknn1,drop=TRUE]

#25% data - testing  for variable y
dt_testsample2 <- BinaryClassifier_dfknn[-BinaryClassifier_dfknn1,drop=TRUE]

#Build nearest neighbors model using k values of 32

modelbinarydfknn <- knn(data.frame(dt_Training1), data.frame(dt_testsample1), cl=dt_Training1, k=32)

#Calculate the proportion of correct classification for k = 32
Accuracyofmodelknn <- 100 * sum(dt_Training2 == modelbinarydfknn)/NROW(dt_testsample2)

```



## Conclusion

Based on my reading and the resutls, Logistic regression classifier different from that of the nearest neighbors. KNN is much accurate than Logistic Regressions.


# References
1.  Calculating UAC: The Area Under a ROC Curve
2.  https://rpubs.com/melonda/190569
3.  https://towardsdatascience.com/binary-classification-with-logistic-regression-31b5a25693c4
4.  https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/str#:~:text=Compactly%20Display%20the%20Str        cture%20of%20an%20Arbitrary%20R%20Object&text=The%20idea%20is%20to%20give,.)%20%2C%20see%20the%20examples.
5.  https://towardsdatascience.com/real-world-implementation-of-logistic-regression-5136cefb8125
6.  http://rstudio-pubs-static.s3.amazonaws.com/514195_93183a1fcf2d4a41b5eeb72f91048815.html
7.  https://www.youtube.com/watch?v=C4N3_XJJ-jU
8.  https://www.datacamp.com/community/tutorials/logistic-regression-R
9. https://rpubs.com/shirokaner/logistic
10. https://rstudio-pubs-static.s3.amazonaws.com/304133_610e9a56b9f7477791b2c1fe3cd8dd42.html
11. https://lgatto.github.io/IntroMachineLearningWithR/supervised-learning.html