---
title: 'Assignment 8.2 : Introduction to machine Learning'
author: "Ragunath Gunasekaran"
date: "July 21 2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---


## Footnote
This is a Footnote test.

## Citations

* R for Everyone
* Discovering Statistics Using R

##  Plot the data from each dataset using a scatter plot

```{r dataframe code, include=FALSE}
binary_data <- read.csv("binary-classifier-data.csv")
trinary_data <- read.csv("trinary-classifier-data.csv")
```


##  Scatter plot of binary dataset

```{r plot, echo=FALSE, message=FALSE}
library(ggplot2)
scatterbin <- ggplot(binary_data, aes(x, y))
scatterbinary <- scatterbin + geom_point() + geom_smooth(method = "lm", aes(fill = label), alpha = 0.1) + labs(x = "Value of X", y = "Value of y", colour = "label")

plot(scatterbinary)

```


##  Scatter plot of trinary dataset


```{r plot1, echo=FALSE, message=FALSE}

scattertrin <- ggplot(trinary_data, aes(x, y))
scattertrinary <- scattertrin + geom_point() + geom_smooth(method = "lm", aes(fill = label), alpha = 0.1) + labs(x = "Value of X", y = "Value of y", colour = "label")
plot(scattertrinary)


```

## k nearest neighbors algorithm categorizes
## Fiting a k nearest neighbors model for each dataset for k=3, k=5, k=10, k=15, k=20, and k=25. Compute the accuracy of the resulting models for each value of k. Plot the results in a graph where the x-axis is the different values of k and the y-axis is the accuracy of the model.


```{r dataframe code1, include=FALSE}
# Data Normalization
binary_data1 <- binary_data[, c("x", "y")]
trinary_data1 <- trinary_data[, c("x", "y")]

# Create train and test datasets for binary data
set.seed(111)
dat.binary <- sample(1:nrow(binary_data1),size=nrow(binary_data1)*0.7,replace = FALSE) #random selection of 70% data.
 
train.binary <- binary_data[dat.binary,] # 70% training data
test.binary <- binary_data[-dat.binary,] # remaining 30% test data


# Create train and test datasets for trinary data
set.seed(123)
dat.d <- sample(1:nrow(trinary_data1),size=nrow(trinary_data1)*0.7,replace = FALSE) #random selection of 70% data.
 
train.trinary <- trinary_data[dat.d,] # 70% training data
test.trinary <- trinary_data[-dat.d,] # remaining 30% test data


train.binarylabels <- binary_data[dat.binary,1,drop=TRUE]
test.binarylabels <-binary_data[-dat.binary,1,drop=TRUE]

```
```{r plot2, echo=FALSE, message=FALSE}
##  Find the number of observations in binary train data
NROW(train.binary) 

#  Find the number of observations in binary test data
NROW(test.binary)

#  Find the number of observations in binary train data
NROW(train.binarylabels)

#  Find the number of observations in binary test data
NROW(test.binarylabels)



#Build models with different K values for binary data
#Store K values and accuracy values in a dataframe for different k values
library(class)
list2 <- list(3,5,10,15,20,25)
i=1
k.binaryaccur=1
for (i in list2){
    knn.mod <- knn(train=train.binary, test=test.binary, cl=train.binarylabels, k=i)
    k.binaryaccur[i] <- 100 * sum(test.binarylabels == knn.mod)/NROW(test.binarylabels)
    k=i
    cat(k,'=',k.binaryaccur[i],'')
    }
    

#Creating seperate dataframe for trinary label data
train.trinarylabels <- trinary_data[dat.d,1,drop=TRUE]
test.trinarylabels <-trinary_data[-dat.d,1,drop=TRUE]

# Find the number of observations in trinary train data
NROW(train.trinary) 

#  Find the number of observations in trinary test data
NROW(test.trinary)

##  Find the number of observations in trinary train data
NROW(train.trinarylabels) 

#  Find the number of observations in trinary test data
NROW(test.trinarylabels)

#Store K values and accuracy values in a dataframe for different k values
library(class)
list3 <- list(3,5,10,15,20,25)
j=1
k.trinaryaccur=1
for (j in list3){
    knn.modtri <- knn(train=train.trinary, test=test.trinary, cl=train.trinarylabels, k=j)
    k.trinaryaccur[j] <- 100 * sum(test.trinarylabels == knn.modtri)/NROW(test.trinarylabels)
    k=j
    cat(k,'=',k.trinaryaccur[j],'')
    }
    
    
#  Plot Accuracy values with k values

plot(k.trinaryaccur, type="b", xlab="K- Value",ylab="Accuracy level")


#  Plot Accuracy values with k values
plot(k.binaryaccur, type="b", xlab="K- Value",ylab="Accuracy level")
```

# References
1. Bernard Marr. (2016). Supervised V Unsupervised Machine Learning – What’s The Difference?
2. Bernard Marr. (2016). What Is The Difference Between Artificial Intelligence And Machine Learning?
3. Bernard Marr. (2016). What Is The Difference Between Deep Learning, Machine Learning and AI?